Spatial Reasoning and Statistical Graphics

### 2014-10-20 Paper/Draft1/TestingVisualAptitude.Rnw
Unscaled answers

\subsection{Scoring and Scaling of Answers}\label{sec:scaling}

@@

\paragraph{Variance consideration of scaled test score:}
Let $X_{n, k}$ be a participant's (unscaled) score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
\begin{eqnarray*}
\text{Var}(X_{n, k}) &=& n^2 \text{Var}(X_{1, k}) = \\
&=& n^2 (\underbrace{1/k \cdot 1^2}_{\text{correct answer}}+ \underbrace{(-1/(k-1))^2 \cdot (k-1)/k}_{\text{wrong answer}}) = \\
&=& n^2/(k-1).
\end{eqnarray*}
\newtext{The scaled test score $Z_{n,k}$ is $1/n \sqrt{(k-1)} X_{n,k}$ and the statement for the variance of $Z_{n,k}$ follows directly.}

The above consideration only assumes independence between questions, which is reasonable. While we only consider a test consisting of questions with the same number of choices $k$, an extension to varied number of answers is trivial and has been done in the adjustment for the figure classification score.

To facilitate comparison of the visuospatial skills tests utilized in this study with the lineups, we will scale the test results by a factor of $1/n \sqrt{k-1}$, where $n$ is the number of questions in a test and $k$ is the number of offered answers on each questions (out of which only one is correct). This scaling ensures that test scores across different types are comparable. All test scores are now normalized such that they have under the assumption of random guessing an expected value of zero and a standard deviation of 1. In particular, this leads to a theoretical range of a normalized test score $Z_{n, k}$ of $[-\sqrt{k-1}, \sqrt{k-1}]$.

@@

\comment{\textit{We may want to include a table of scaled scores here, so that we don't just bust right into lineup discussion without contextualizing everything properly.}}
\newtext{Applying the scaling protocol discussed in section \ref{sec:scaling}, we see that the ranges of scaled lineup and visuospatial test scores do not include zero; this indicates that we do not see random guessing from participants in any task. Figure \ref{fig:ScaledScores} shows the range of possible scores and the observed score distribution. } 
<<ResultsScaledScores, echo=F, include=T, fig.cap="Scaled test scores for lineups and visuospatial tests. As none of the participants scored at or below zero, we can conclude that there is little evidence of random guessing. \\label{fig:ScaledScores}", fig.width=8, fig.height=3, out.width='.9\\textwidth'>>=

>>>>

\subsection{Test Scoring}\label{sec:scaling}

@@

% 
% \paragraph{Variance consideration of scaled test score:}
% Let $X_{n, k}$ be a participant's (unscaled) score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
% \begin{eqnarray*}
% \text{Var}(X_{n, k}) &=& n^2 \text{Var}(X_{1, k}) = \\
% &=& n^2 (\underbrace{1/k \cdot 1^2}_{\text{correct answer}}+ \underbrace{(-1/(k-1))^2 \cdot (k-1)/k}_{\text{wrong answer}}) = \\
% &=& n^2/(k-1).
% \end{eqnarray*}
% \newtext{The scaled test score $Z_{n,k}$ is $1/n \sqrt{(k-1)} X_{n,k}$ and the statement for the variance of $Z_{n,k}$ follows directly.}
% 
% The above consideration only assumes independence between questions, which is reasonable. While we only consider a test consisting of questions with the same number of choices $k$, an extension to varied number of answers is trivial and has been done in the adjustment for the figure classification score. 
% 
% To facilitate comparison of the visuospatial skills tests utilized in this study with the lineups, we will scale the test results by a factor of $1/n \sqrt{k-1}$, where $n$ is the number of questions in a test and $k$ is the number of offered answers on each questions (out of which only one is correct). This scaling ensures that test scores across different types are comparable. All test scores are now normalized such that they have under the assumption of random guessing an expected value of zero and a standard deviation of 1. In particular, this leads to a theoretical range of a normalized test score $Z_{n, k}$ of $[-\sqrt{k-1}, \sqrt{k-1}]$.

@@

\comment{\textit{We may want to include a table of scaled scores here, so that we don't just bust right into lineup discussion without contextualizing everything properly.}}
\newtext{Applying the grading protocol discussed in section \ref{sec:scaling}, we see that the ranges of lineup and visuospatial test scores do not include zero; this indicates that we do not see random guessing from participants in any task. Figure \ref{fig:Scores} shows the range of possible scores and the observed score distribution. } 
<<ResultsScaledScores, echo=F, include=T, fig.cap="Test scores for lineups and visuospatial tests. As none of the participants scored at or below zero, we can conclude that there is little evidence of random guessing. \\label{fig:Scores}", fig.width=8, fig.height=3, out.width='.9\\textwidth'>>=
